Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-14 chi2012.tex
added more in results section

The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web.  The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.   We describe our methodology below.

In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.  For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).  We analysed the data collected from this study using a thematic approach.  From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).  We performed a statistical analysis of the features of the available data.  In the next two sections, we describe our two pre-studies.
%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.  In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one. This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted. The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined. In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.  Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.

@@

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=18cm]{img/dragdrop}
\caption{WebBox uses a Drag and Drop interface to add data into the workspace.}
\label{fig:dragdrop}
\end{center}
\end{figure*}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=8cm]{img/multipathing}
\caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
\label{fig:multipathing}
\end{center}
\end{figure}

\begin{figure*}[htbp]
\begin{center}
\includegraphics[width=18cm]{img/sameas}
\caption{When two instances represent the same entity, users can drag one on top of the other, and they are combined.}
\label{fig:sameas}
\end{center}
\end{figure*}

@@

\subsection{Participants}
We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.  All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.  Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.  All of the participants had previously encountered tasks requiring them to use and evaluate data from multiple websites.

@@

\begin{table*}[t]
\caption{Tasks, choices and data sources}
\begin{center}

\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
\hline
Task	 &Choices	&Data Sources\\
\hline
1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
\hline
2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
\hline
3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant. & www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
\hline
4) Restaurant they would book for 12 friends in Glagow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollopes Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia. & www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
\hline
\end{tabular}
\end{center}
\label{tab:studyfactors}
\end{table*}%

Each task had an A and B condition.  The A condition, the participant could only use WebBox to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or spreadsheets in Microsoft Excel containing the data taken from the data sources to make their choice.

@@

\subsection{Data Collection}
Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.  During the lab study, we recorded the audio and the participant's actions on screen. We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.

@@

% behaviours in websites

Users who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In
particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of
university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly
in searching for university courses and course requirements. Similarly, participants found it difficult to understand the
locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto
the same map using the websites. Overall, the users that spent the most time using websites exhibited significant
frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations),
and all ended up falling back to using the supplied spreadsheets.

% behaviours in excel
One of the designs of the study was to produce a spreadsheet worksheet for each website, which meant that users
had to switch between the worksheets to gather data from multiple sources. Different approaches to using the spreadsheets
were observed. Some users were keen to collate all data onto a single sheet, so that they could compare the values
all at once. However, this approach was error prone, for example participants noticing too late that the ordering of rows
was different on each sheet, so that their pasted values were incorrectly mapped. Others used pen and paper to collate
the values, which was often faster and resulted in less observed frustration. While the majority of users did not
attempt anything complex with Excel (in fact only one users used charts), there was some advanced use of Excel observed.
One user explained that they did not like to look at numbers, and so coloured the cells of the types of restaurant, and the
friends' preferences, and matched the colours to make their decision. Another user calculated their own weighted metric
of which university to choose by averaging different metrics supplied by the data sources.

% webbox client

The most commonly observed use of WebBox was to collate the data on-screen, and compare multiple values directly.
About half of the observed users actually used the instance combination feature of WebBox, with the other half able
to successfully make decisions without using it. Those who tried to use it were largely successful, and were able
to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened
up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result
of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.

%% more from doc here

% weird and wonderful
A number of users admittedly introduced their own bias into the choices: ``I prefer Indian food, so I will choose the Indian
restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating
those that required bus transfers (the protocol did not mention routing/transfers at all).

>>>>

In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.  For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a tasks requiring multiple websites (see Section \ref{sec:prestudy1}).  We analysed the data collected from this study using a thematic approach.  From our first pre-study's qualitative analysis, we designed and performed our second study which analyses existing on data available from APIs and structured data feeds on the Web (see Section \ref{sec:prestudy2}).  The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.  We performed a statistical analysis of the features of the available data.  In the next two sections, we describe our two pre-studies.
%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.  In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one. This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted. The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined. In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.  Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.

@@

@@

\subsection{Participants}
We recruited participants by calling for participant on the University of Southampton's international student mailing list.  There is no guarantee of the mailing list's recipients first language, however all of them passed the university English proficiency test.  We required that all participants had encountered tasks that required them to use and evaluate data from multiple websites.

@@

\begin{table*}[t]
\caption{Tasks, choices and data sources}
\begin{center}
\small
\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
\hline
Task	 &Choices	&Data Sources\\
\hline
1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
\hline
2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
\hline
3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant. & www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
\hline
4) Restaurant they would book for 12 friends in Glagow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollopes Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia. & www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
\hline
\end{tabular}
\end{center}
\label{tab:studyfactors}
\normalsize
\end{table*}%

Each task had an A and B condition.  The A condition, the participant could only use WebBox to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or workbook in Microsoft Excel containing the data taken from the data sources to make their choice on individual spreadsheets.

@@

\subsection{Data Collection}
Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.  During the lab study, we recorded the audio and the participant's actions on screen. We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.  At the end of the study the participant completed an exit survey.

@@

We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.  All of the participants spoke English, and for 12 of them, English was not their first language.  Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.

The information in Table \ref{} was captured during the studies, so that we could statistically analyses the features used by the participants.

\begin{table*}[htdp]
\caption{default}
\small
\begin{center}
\begin{tabular}{|r|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
\hline
Participants & 1 &2&3&4&5&6&7&8&9&10&11&12&13&14&15&16&17&18&19&20\\
\hline\hline
gender&f&m&f&f&f&f&f&m&f&f&m&f&m&f&m&m&m&m&m&m\\
A condition first task &y&y&y&n&n&y&n&n&y&y&n&n&y&y&n&n&n&n&n&n\\
\hline\hline
A Condition: task &4&2&3&4&3&&2&1&4&1&4&1&3&2&3&2&1&4&3&2\\
paper used&n&n&n&n&n&n&y&n&y&y&n&y&n&n&y&n&n&n&n&n\\
charts used &1&2&2&0&2&4&5&0&0&0&2&3&2&0&0&0&3&3&4&1\\
maps used &1&1&2&2&2&1&1&0&1&0&1&1&2&1&0&1&0&0&2&1\\ 
same as &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
number data sources  &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, 1.3, 1.2& 4.2& 1.4, 1.5, 1.3& 3.2& 2.1, 2.3, 2.6& 3.5, 3.3&2.6, 2.3, 2.5&1.6, 1.5, 1.3&4.1&3.3&2.4, 2.5, 2.6\\
\hline\hline
 B Condition:task &3&&4&3&4&2&1&2&2&3&2&3&1&4&1&4&2&2&1&3\\
 paper used &n&n&n&n&n&n&y&n&y&y&n&y&n&y&y&n&n&y&n&n\\
 spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
 number of spreadsheets used&5/5&4/4&2/5&1.5&2/5&5/5&4/4&2/4&3/4&&2/4&4/5&2/4&4/5&2/4&2/5&4/4&4/4&0&5/5\\
 number of websites used&1&1&4&3&3&0&2&3&2&&1&2&3&1&2&3&2&0&4&0\\
 chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4,1.2,1.6&2.1&2.3,2.6,2.5&3.5&2.2,2.3,2.6&3.5&1.1&4.1&1.6&4.2&2.6,2.3, 2.5&2.3,2.5,2.6&1.1, 1.3, 1.4&3.5\\
 \hline
\end{tabular}
\end{center}
\label{default}
\normalsize
\end{table*}

In the A condition, where participants used WebBox to complete their tasks they used it to collate the data on-screen, and compare multiple values directly. About half of the observed users actually used the instance combination feature of WebBox, with the other half able to successfully make decisions without using it. Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.

In the B conditions, users were given the option of using websites and or a spreadsheet to make their choice.  Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly in searching for university courses and course requirements. Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites. Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.

Participants who opted to use the spreadsheet to support their decisions switched between worksheets to view data from different websites.  They used different approaches to using the spreadsheets, some participants were keen to collate all data onto a single sheet, so that they could compare the values all at once. However, this approach was error prone, for example participants noticing too late that the ordering of rows was different on each sheet, so that their pasted values were incorrectly mapped. Others used pen and paper to collate the values, which was often faster and resulted in less observed frustration. While the majority of users did not
attempt anything complex with Excel (in fact only one users used charts), there was some advanced use of Excel observed.
One user explained that they did not like to look at numbers, and so coloured the cells of the types of restaurant, and the
friends' preferences, and matched the colours to make their decision. Another user calculated their own weighted metric
of which university to choose by averaging different metrics supplied by the data sources.

Notably, a number of users admittedly introduced their own bias into the choices in both A and B conditions: ``I prefer Indian food, so I will choose the Indian restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating those that required bus transfers (the protocol did not mention routing/transfers at all).

In the exit survey, the users noted that they would use WebBox for finding a property to rent/buy, choosing a job and purchasing electronic devices, in the future.  However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the spontaneity, but would use it for making more important decisions.  In order to improve our solutions the participants suggested making the origin of the information clearer, sorting instances, and explanations of ratings and scoring systems (however some of these are not transparent on websites).
